Atmospheric escape and star-planet-interaction in the HAT-P-67 b system

### 2023-02-08 paper/paper1/ms.tex
Greater depth at ingress

The out-of-transit excess exhibits approximately the same line width as the stellar lines, which are rotationally broadened by a $v\sin{i}$ of 30-35 km/s \citep{2017AJ....153..211Z}.  The large size of the gas stream overfills the transit chord, which means both the advancing blue side and receding red side of the rotating star are nearly continuously obscured in this spin-orbit-aligned system \citep{2017AJ....153..211Z}. The planet transits near the Stellar equator \citep{2017AJ....153..211Z}, meaning the transit chord covers about 10$\%$ of the projected stellar surface area.  Hypothetically a nearly 100$\%$ opaque optically thick stream of gas covering only the transit chord could reproduce the observed 10$\%$ absorption excess.  Alternatively---and more likely---an optically thin gas stream has some additional vertical extent perpendicular to the plane of orbit, which would act to overfill the entire projected stellar disk.  Assume the gas has a spatially uniform optical depth of about 0.1 would reproduce the 10\% absorption excess.

The stream must have a finite vertical extent

>>>>

The out-of-transit excess exhibits approximately the same line width as the stellar lines, which are rotationally broadened by a $v\sin{i}$ of 30-35 km/s \citep{2017AJ....153..211Z}.  The large size of the gas stream overfills the transit chord, which means both the advancing blue side and receding red side of the rotating star are nearly continuously obscured in this spin-orbit-aligned system \citep{2017AJ....153..211Z}. The planet transits near the Stellar equator \citep{2017AJ....153..211Z}, meaning the transit chord covers about 10$\%$ of the projected stellar surface area.  Hypothetically a nearly 100$\%$ opaque optically thick stream of gas covering only the transit chord could reproduce the observed 10$\%$ absorption excess.  Alternatively---and more likely---an optically thin gas stream has some additional vertical extent perpendicular to the plane of orbit, which would act to overfill the entire projected stellar disk.  Assuming the gas has a spatially uniform optical depth of about 0.1 would reproduce the 10\% absorption excess.

The excess absorption appears deepest and with the narrowest line width at the moment of ingress.  Gas overfilling the stellar disk does not contribute to the excess absorption, and so the ratio of the vertical component of the intial velocity of the gas (perpendicular to the orbital plane) to the Keplerian shear controls how much time a given gas parcel has before it leaves the projected orbital chord.
